---
title: "R Notebook"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
# 1.
setwd("C:/Users/Avinav/UMD/Sem 2/BUDT758T/Assignment 1/")

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
```

```{r}

# 2.
library("gapminder")
dfGap <- gapminder

```

```{r}

# 3.a
skim(dfGap)

```

```{r}

# 3.b
dfGap %>% 
  filter(year == 2007) %>% 
  arrange(desc(lifeExp))

```

```{r}

# 3.c
dfGapNew <- dfGap %>% 
  group_by(country) %>% 
  mutate(totalGDP = pop*gdpPercap) %>% 
  filter(year == 2007) %>% 
  ungroup() %>%  
  arrange(desc(totalGDP)) 

dfGapNew

```

```{r}

# 3.c.iii
dfGap %>% 
  filter(year == 2007) %>% 
  select(country, gdpPercap) %>% 
  arrange(desc(gdpPercap))

```

```{r}

# 3.d
dfGapNew %>%  
  group_by(continent) %>% 
  summarise(mdnLifeExp = median(lifeExp), mdnTotGDP = median(totalGDP)) %>%
  ungroup() %>% 
  arrange(desc(mdnLifeExp))

```

```{r}

# 4.a.i
dfGapNew %>% 
  ggplot() + geom_point(aes(x = totalGDP, y = lifeExp))

```

```{r}

# 4.a.ii
dfGapNew %>% 
  filter(totalGDP > 1e+12) %>% 
  ggplot() + geom_point(aes(x = totalGDP, y = lifeExp))

```

```{r}

# 4.a.iii
dfGapNew %>% 
  filter(totalGDP > 1e+12) %>% 
  ggplot(aes(x = totalGDP, y = lifeExp)) + 
  geom_point() + 
  geom_label(aes(label = country))

```

```{r}

# 4.a.iv
library("ggrepel")

```
```{r}

dfGapNew %>% 
  filter(totalGDP > 1e+12) %>% 
  ggplot(aes(x = totalGDP, y = lifeExp)) + 
  geom_point() + 
  geom_label_repel(aes(label = country))

```

```{r}

# 4.a.v
dfGapNew %>% 
  filter(totalGDP > 1e+12) %>% 
  ggplot(aes(x = totalGDP, y = lifeExp, color = continent)) + 
  geom_point() + 
  geom_label_repel(aes(label = country))

```

```{r}

# 4.b
dfGap %>% 
  filter(year == 2007) %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_smooth()

```

```{r}

# 4.c
boxPlotsForAll <- dfGap %>% 
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot() +
  labs(title = "Life expectancy across continents")
boxPlotsForAll

```

```{r}

# 4.c.iii
ggplotly(boxPlotsForAll)

```

```{r}

# 4.d
gdpOverTime <- dfGap %>%  
  group_by(continent, year) %>% 
  summarise(mdnGDPPercap = median(gdpPercap)) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = mdnGDPPercap, color = continent)) +
  geom_line() +
  labs(x = "Year", y = "Median GDP per capita")

gdpOverTime
  

```

```{r}

# 4.d.ii
dfGap %>%  
  group_by(continent, year) %>% 
  summarise(meanGDPPercap = mean(gdpPercap)) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = meanGDPPercap, color = continent)) +
  geom_line() +
  labs(x = "Year", y = "Mean GDP per capita")

```

```{r}

# 4.d.iii
ggplotly(gdpOverTime)

```

